Jaigo Technical Report

### 2007-12-19 trunk/paper/paper.tex
fix up LaTeX and don't version build output

The exploration of alternative future board states often yields duplicate (potential) states; it is helpful to memoize board evaluation such that a board state may be evaluated only once, regardless of the paths by which it is reached \cite{Zobrist1970b}. The Zobrist hash function is useful in this capacity: (reduce xor (map f board-positions)), f: (x, y, color) -> Z.

@@

    * Finding a good evaluation function is hard. The human heuristics are vague, and when/how they get applied is even more vague. Zobrist divided the game into phases for the purpose of selecting evaluation functions \cite{Zobrist1970b}. Ryder distinguished between tactical and strategic utility \cite{Burmeister}.
    * fight between optimizing the main loop (evaluating the board) vs optimizing the small loops (given a board evaluation, local tricks to improve evaluation function). GNUGO v Go++
    * GA \& NN seemed promising at first, but no commercial package uses them due to lack of speed.
    * If all old techniques are inadequate and new techniques don't pan out, whither computer go?

Our project seeks to create a JavaScript-based Go-engine. JavaScript is a dynamically typed, imperative language that supports the functional and object-oriented paradigms \cite{ECMA-262}. It is ubiquitous due to its inclusion in nearly every Web browser, and recently has enjoyed a surge of popularity among programmers due to the availability of cross-platform libraries and the success of applications such as GMail. As professional programmers, we viewed this project as an opportunity to explore not only AI but also renewed interest in the JavaScript language. Also, JavaScript is the only supported means of developing iPhone applications. iPhone is an intrinsically interesting platform; its communication capabilities and computational limitations only heighten the interest.

@@

SimpleGo v1

50kyu greedy search (random move from list of moves with worst opponent score)

SimpleGo v3

30kyu $\alpha$-$\beta$ pruned search

Crawler

?? (~30kyu) maintain line strength/shape

\chapter{Technical details}
Code Listing
- Implementation of Benson's algorithm for determining unconditional life
- Greedy Search algorithm: select\_scored\_move
- Crawler algorithm: select\_crawler\_move

%(5)    Discussions, observations, and comparisons.
%(6)    Recommendations, especially for future work and unsolved issues.
%(7)    References.

>>>>

The exploration of alternative future board states often yields duplicate (potential) states; it is helpful to memorize board evaluation such that a board state may be evaluated only once, regardless of the paths by which it is reached \cite{Zobrist1970b}. The Zobrist hash function is useful in this capacity: (reduce xor (map f board-positions)), f: (x, y, color) -> Z.

There are traditional concepts in go of blocks and eyes. Blocks are simply contiguously-connected stones of the same color. Eyes are empty spaces which are completely surrounded by stones of the same color. A good implementation of computer Go will represent these concepts, as they provide vital information when generating moves, evaluating the state of the board, and in scoring the board.

@@

There are several significant issues to consider when programming computer Go. First, it is difficult to find an appropriate evaluation function. The human heuristics are vague, and the parameters for their application  are even more vague. Zobrist divided the game into phases for the purpose of selecting evaluation functions \cite{Zobrist1970b}. On the other hand, Ryder distinguished between tactical and strategic utility \cite{Burmeister}.

Second, there is a fight between optimizing the main loop -- that is, evaluating the board -- versus optimizing the small loops (given a board evaluation, local tricks to improve evaluation function). There is yet no standard, robust board evaluation algorithm. If a main board evaluation loop is established for playing computer Go and optimizations are made to the smaller evaluation functions within the loop, this would change the way the board is evaluated.  However, though such developments would potentially better the speed, completeness, or accuracy of the evaluation, they may also potentially damage the smaller evaluation functions.

Third, while GA \& NN initially seemed promising, no commercial package uses them due to lack of speed.

Thus, if all older techniques are inadequate and new techniques do not pan out, whither computer Go? Our project seeks to create a JavaScript-based Go-engine. JavaScript is a dynamically typed, imperative language that supports the functional and object-oriented paradigms \cite{ECMA-262}. It is ubiquitous due to its inclusion in nearly every Web browser, and recently has enjoyed a surge of popularity among programmers due to the availability of cross-platform libraries and the success of applications such as GMail. As professional programmers, we viewed this project as an opportunity to explore not only AI but also renewed interest in the JavaScript language. Also, JavaScript is the only supported means of developing iPhone applications. iPhone is an intrinsically interesting platform; its communication capabilities and computational limitations only heighten the interest.

@@

We chose to implement two different move choice engines, each of which provides a different playing experience. SimpleGo v1, a 50kyu greedy search move generator which picks a random move from list of moves with worst opponent score. Crawler, an unranked move generator which analyzes the current blocks on the board and picks the move that both connects to a block and results in the largest number of liberties for the block.

\chapter{Technical details}
Code Listing
- Implementation of Benson's algorithm for determining unconditional life: Analyze\_Color\_Unconditional\_Status
- Greedy Search algorithm: select\_scored\_move
- Crawler algorithm: select\_crawler\_move

\chapter{Discussions, observations, and comparisons}
We implemented two move choice engines based on two different algorithms, scoring for color and blocks based on Benson's algorithm, and representations of the board, the game, eyes, blocks, and numerous language-specific utilities. Our goals were compactness and speed due to the target platform limitations.

Due to issues with the jaigo board representation, our measurements were limited to the first 100 moves.

Regarding compactness, the total size of all jaigo files, including images, scripts, and html, was 102221 bytes for our initial release. The full implementation of the jaigo code is 66071 bytes \cite{SpeedReport}.

Regarding speed, we do not anticipate that the choice of processor and platform of any modern PC will have a significant change to the performance of jaigo.

For the first 50 moves on a 19x19 board with human (black) versus crawler (white),
\begin{tabular}{ccccccccc}
\hline
Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\
\hline
(no name) & 109675 & 25.68 & 14310.594ms &	36908.812ms &	0.337ms &	0.059ms	& 1593.542ms	& simple\_go.js (line 407)\\
\hline
flood\_mark & 1618 & 14.17 & 7897.73ms & 25481.245ms & 15.749ms & 0.162ms & 1635.18ms & simple\_go.js (line 807)\\
\hline
(no name) &	194678 &	11.75	& 6551.626ms &	6554.733ms &	0.034ms &	0.008ms &	1555.243ms &	simple\_go.js (line 819)\\
\hline
string\_as\_move &	56065 &	7.48 &	4168.485ms &	4169.549ms &	0.074ms &	0.011ms &	1591.844ms &	simple\_go.js (line 192)\\
\hline
iterate	& 109676 &	7.19 &	4005.749ms &	22595.463ms &	0.206ms &	0.011ms &	1593.499ms &	simple\_go.js (line 391)\\
\hline
(no name) &	283642 &	4.9 &	2732.499ms &	18589.343ms &	0.066ms &	0.015ms &	1593.144ms &	simple\_go.js (line 416)\\
\hline
iterateOwnProperties &	5811	& 4.38	& 2440.378ms &	2847.852ms &	0.49ms &	0.004ms &	2.917ms &	simple\_go.js (line 132)\\
\hline
Set\_Goban &	2150 &	3.53 &	1968.651ms &	1969.095ms &	0.916ms &	0.004ms &	942.058ms &	simple\_go.js (line 528)\\
\hline
(no name) &	42734 &	2.99 &	1666.478ms &	1666.716ms &	0.039ms &	0.008ms &	935.83ms &	simple\_go.js (line 835)\\
\hline
update &	3662 &	2.9 &	1613.61ms &	1614.099ms &	0.441ms &	0.004ms &	1560.416ms &	simple\_go.js (line 117)\\
\hline
Add\_Block &	2340 &	2.8 &	1559.92ms &	1560.774ms &	0.667ms &	0.007ms &	1529.63ms &	simple\_go.js (line 757)\\
\hline
Calculate\_Neighbour &	448	2.67 &	1487.337ms &	12205.939ms &	27.245ms &	0.019ms &	1579.849ms &	simple\_go.js (line 829)\\
\hline
Simple\_Same\_Block &	2149 &	1.93 &	1073.573ms &	8185.472ms &	3.809ms &	0.002ms &	1556.376ms &	simple\_go.js (line 606)\\
\hline
List\_Block\_Liberties &	3210 &	1.56 &	871.609ms &	5057.619ms &	1.576ms &	0.012ms &	1593.058ms &	simple\_go.js (line 863)\\
\hline
\end{tabular}

For the first 100 moves of a game on a 19x19 board, with crawler (black) versus crawler (white),

\begin{tabular}{ccccccccc}
	\hline
	Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\	
\hline
(no name) &	404494 &	32.8 &	71255.203ms &	147518.241ms &	0.365ms &	0.059ms &	1650.149ms &	simple\_go.js (line 407)\\
\hline
flood\_mark &	3838 &	15.94 &	34634.022ms &	96713.52ms &	25.199ms &	0.206ms &	1736.537ms &	simple\_go.js (line 807)\\
\hline
iterate &	404495 &	8.75 &	19004.696ms &	76254.33ms &	0.189ms &	0.011ms &	1650.101ms &	simple\_go.js (line 391)\\
\hline
(no name) &	735190 &	8.72 &	18941.197ms &	18941.987ms &	0.026ms &	0.008ms &	1572.648ms &	simple\_go.js (line 819)\\
\hline
string\_as\_move &	242765 &	5.86 &	12724.373ms	& 12725.279ms &	0.052ms &	0.011ms &	1629.186ms &	simple\_go.js (line 192)\\
\hline
(no name) &	1041519 &	4.7 &	10209.513ms &	57248.372ms &	0.055ms &	0.015ms &	1649.869ms &	simple\_go.js (line 416)\\
\hline
List\_Block\_Liberties &	7706 &	4.02 &	8727.331ms &	21457.809ms &	2.785ms &	0.013ms &	1630.607ms &	simple\_go.js (line 863)\\
\hline
iterateOwnProperties &	14752 &	3.68 &	7985.019ms &	9340.503ms &	0.633ms &	0.004ms &	2.419ms &	simple\_go.js (line 132)\\
\hline
update &	8742 &	3.18 &	6911.102ms &	6911.93ms &	0.791ms &	0.004ms &	1648.007ms &	simple\_go.js (line 117)\\
\hline
Calculate\_Neighbour &	661 &	3.03 &	6593.032ms &	60385.478ms &	91.355ms &	0.115ms &	1679.244ms &	simple\_go.js (line 829)\\
\hline
(no name) &	160414 &	2.51 &	5444.827ms &	5445.809ms &	0.034ms	& 0.008ms &	944.464ms &	simple\_go.js (line 835)\\
\hline
Add\_Stone &	6010 &	1.1 &	2392.214ms	& 191461.602ms	& 31.857ms &	1.524ms &	1785.447ms &	simple\_go.js (line 641)\\
\hline
\end{tabular}

Median move response time for 10 games into the first 50 moves was was 1 second, with the high being 3 seconds and the low being unmeasurable.

On an iPhone over 54mbps 802.11g wifi with a full signal connected to a 3mbps cable modem, the median jaigo load time was 5 seconds, with the initial load time being the longest measurement at 15 seconds. On an iPhone over AT\&T 2.5G EDGE with a full signal, the median jaigo load time was 21 seconds, with a low of 17 seconds and a high of 30 seconds.

Median move times in other go engines (gnugo especially)?

\chapter{Recommendations, especially for future work and unsolved issues}
We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo. However, we have not quantified this claim.

The performance of $\alpha-\beta$ search pruning in javascript should be investigated. like SimpleGo v3, a 30kyu $\alpha$-$\beta$ pruned search move generator.

There is much more optimization and correctness work to be done to the jaigo board and game representations. For example, the move speed of jaigo could be measured on many diverse computing platforms. Jaigo could be also used as a "widget", a popular way of packaging small javascript and html applications into a format that is used by many different web sites and operating systems such as Windows Vista, Mac OS X, Facebook, Google, and others.

There are many open issues being tracked by the jaigo project at this time. Addressing them may lead to better, more correct performance. We anticipate continuing work on this project.

%(7)    References.
